---
title: "Data621 Assignment3"
author: "Ritesh Lohiya"
date: "June 30, 2018"
output: html_document
---

#Overview
objective is to build a binary logistic regression model on the training data set to predict whether the neighborhood will be at risk for high crime levels. Below is a short description of the variables in the dataset.

zn: proportion of residential land zoned for large lots (over 25000 square feet)

indus: proportion of non-retail business acres per suburb

chas: a dummy var. for whether the suburb borders the Charles River (1) or not (0)

nox: nitrogen oxides concentration (parts per 10 million)

rm: average number of rooms per dwelling

age: proportion of owner-occupied units built prior to 1940

dis: weighted mean of distances to five Boston employment centers

rad: index of accessibility to radial highways

tax: full-value property-tax rate per $10,000

ptratio: pupil-teacher ratio by town

black: 1000 $(B_k - 0.63)^2$ where Bk is the proportion of blacks by town

lstat: lower status of the population (percent)

medv: median value of owner-occupied homes in $1000s

target: whether the crime rate is above the median crime rate (1) or not (0) (response variable)

```{r}
library(readr)
library(kableExtra)
library(tidyverse)
library(knitr)
library(psych)
library(gridExtra)
library(usdm)
library(mice)
library(ggiraph)
library(cowplot)
library(reshape2)
library(corrgram)
library(caTools)
library(caret)
library(ROCR)
library(pROC)
library(reshape2)
library(Amelia)
library(qqplotr)
library(moments)
library(car)
library(MASS)
library(geoR)
```

#DATA EXPLORATION:

```{r}
crime_train <- read.csv("https://raw.githubusercontent.com/Riteshlohiya/Data621-Week3-Assignment3/master/crime-training-data.csv") 
crime_eval <- read.csv("https://raw.githubusercontent.com/Riteshlohiya/Data621-Week3-Assignment3/master/crime-evaluation-data.csv")
summary(crime_train)
```

Visual Exploration:

Now we will see the missing values in the dataset. For this i have used Amelia package

```{r}
missmap(crime_train, main = "Missing values vs observed",  color='dodgerblue')
```

There are no missing values in the dataset.

Lets now dig into our response variables.

1.	Response Variable zn - proportion of residential land zoned for large lots (over 25000 square feet). We can see there are more zeros values for zn and also has positive skewness. Also there appears to be relationship between crime rates and zn.

```{r}
with(crime_train, c(summary(zn), SD=sd(zn), Skew=skewness(zn), Kurt=kurtosis(zn)))

hist <- ggplot(crime_train, aes(zn)) + geom_histogram(fill = 'dodgerblue', binwidth = 20, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of zn') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=zn)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of zn") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", zn)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of zn', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), zn)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of zn by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

2.	Response Variable: indus - proportion of non-retail business acres per suburb. The histogram below indicates a bi-modal quality to the variable’s distribution, with many values clustering in two ranges.

```{r}
with(crime_train, c(summary(indus), SD=sd(indus), Skew=skewness(indus), Kurt=kurtosis(indus)))

hist <- ggplot(crime_train, aes(indus)) + geom_histogram(fill = 'dodgerblue', binwidth = 5, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of indus') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=indus)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of indus") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", indus)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of indus', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), indus)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of indus by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

3.	Response Variable: chas - a dummy var. for whether the suburb borders the Charles River (1) or not (0). This variable tells us if the neighborhood borders the Charles River (1) or not (0). Close to 7% of the neighborhood borders the Charles River. Of the areas bordering the Charles river 21 are in high crime areas.

```{r}
addmargins(table(crime_train$chas))
addmargins(table(crime_train$chas, crime_train$target))
```

```{r}
ggplot(crime_train, aes(x=target, y=chas)) + geom_jitter(color='seagreen4') + theme_classic() + 
  labs(title ='Jittered Scatter Plot of chas vs.target') + theme(plot.title = element_text(hjust = 0.5)) 
```

4.	Response Variable: nox - nitrogen oxides concentration (parts per 10 million). The variable nox represents the concentration of nitrogen oxide in each Boston area. There is also positive skewness. We also see moderately higher nox variance in high crime areas.

```{r}
with(crime_train, c(summary(nox), SD=sd(nox), Skew=skewness(nox), Kurt=kurtosis(nox)))

hist <- ggplot(crime_train, aes(nox)) + geom_histogram(fill = 'dodgerblue', binwidth = .05, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of nox') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=nox)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of nox") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", nox)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of nox', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), nox)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of nox by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

5. Response Variable: rm -  average number of rooms per dwelling. The predictor rm   is count measure describing the average number of rooms per dwelling. The distridution has heavy tail and has bell curve.

```{r}
with(crime_train, c(summary(rm), SD=sd(rm), Skew=skewness(rm), Kurt=kurtosis(rm)))

hist <- ggplot(crime_train, aes(rm)) + geom_histogram(fill = 'dodgerblue', binwidth = 0.5, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of rm') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=rm)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of rm") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", rm)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of rm', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), rm)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of rm by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

6.	Response Variable: age - proportion of owner-occupied units built prior to 1940. The variable age indicates the proportion of owner occupied units built prior to 1940. This variable has high left skewness. Also there is significantly higher mean percentage of older homes in high crime areas.

```{r}
with(crime_train, c(summary(age), SD=sd(age), Skew=skewness(age), Kurt=kurtosis(age)))

hist <- ggplot(crime_train, aes(age)) + geom_histogram(fill = 'dodgerblue', binwidth = 5, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of age') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=age)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of age") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", age)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of age', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), age)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of age by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

7.	Response Variable: dis - weighted mean of distances to five Boston employment centers. The predictor dist describes the average distance to Boston employment centers. The variable is moderately right skewed. Also we can see that low crime areas are associated with higher average distances to employment centers.

```{r}
with(crime_train, c(summary(dis), SD=sd(dis), Skew=skewness(dis), Kurt=kurtosis(dis)))

hist <- ggplot(crime_train, aes(dis)) + geom_histogram(fill = 'dodgerblue', binwidth = 1, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of dis') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=dis)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of dis") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", dis)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of dis', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), dis)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of dis by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

8.	Response Variable: rad - index of accessibility to radial highways. The rad variable is an integer-valued index measure indicating an area’s accessibility to radial highways. In the boxplots below, there appears to be a significant positive association between high crime rates and rad value.

```{r}
with(crime_train, c(summary(rad), SD=sd(rad), Skew=skewness(rad), Kurt=kurtosis(rad)))

hist <- ggplot(crime_train, aes(rad)) + geom_histogram(fill = 'dodgerblue', binwidth = 1, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of rad') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=rad)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of rad") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", rad)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of rad', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), rad)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of rad by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

9.	Response Variable: tax - full-value property-tax rate per $10,000. The tax variable refers to the the tax rate per $10k of property value. High crime areas also appear to have a strong, positive association with the tax value. This variable is densely distributed around two of the following approximate values: 300 and 700.

```{r}
with(crime_train, c(summary(tax), SD=sd(tax), Skew=skewness(tax), Kurt=kurtosis(tax)))

hist <- ggplot(crime_train, aes(tax)) + geom_histogram(fill = 'dodgerblue', binwidth = 20, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of tax') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=tax)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of tax") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", tax)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of tax', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), tax)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of tax by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

10.	Response Variable: ptratio - ptratio: pupil-teacher ratio by town. The predictor ptratio indicates the average school, pupil-to-student ratio, and has a left skewed distribution. We can see a positive relationship between ptratio and high crime.

```{r}
with(crime_train, c(summary(ptratio), SD=sd(ptratio), Skew=skewness(ptratio), Kurt=kurtosis(ptratio)))

hist <- ggplot(crime_train, aes(ptratio)) + geom_histogram(fill = 'dodgerblue', binwidth = 1, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of ptratio') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=ptratio)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of ptratio") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", ptratio)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of ptratio', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), ptratio)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of ptratio by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

11.	Response Variable: black - 1000 (Bk−0.63)2(Bk−0.63)2 where Bk is the proportion of blacks by town. This variable is heavily left skewed.

```{r}
with(crime_train, c(summary(black), SD=sd(black), Skew=skewness(black), Kurt=kurtosis(black)))

hist <- ggplot(crime_train, aes(black)) + geom_histogram(fill = 'dodgerblue', binwidth = 40, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of black') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=black)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of black") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", black)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of black', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), black)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of black by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

12.	Response Variable: lstat - lower status of the population (percent). The variable lstat indicates the proportion of the population deemed to be of lower status. lstat is right skewed. High crime areas tend to have be associated with larger lstat values.

```{r}
with(crime_train, c(summary(lstat), SD=sd(lstat), Skew=skewness(lstat), Kurt=kurtosis(lstat)))

hist <- ggplot(crime_train, aes(lstat)) + geom_histogram(fill = 'dodgerblue', binwidth = 2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of lstat') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=lstat)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of lstat") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", lstat)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of lstat', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), lstat)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of lstat by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

13.	Response Variable: medv - median value of owner-occupied homes in $1000s. The median value of residential homes in a given area. The variable is slightly right skewed, and high values of medv appear to be associated with lower crime rates. 

```{r}
with(crime_train, c(summary(medv), SD=sd(medv), Skew=skewness(medv), Kurt=kurtosis(medv)))

hist <- ggplot(crime_train, aes(medv)) + geom_histogram(fill = 'dodgerblue', binwidth = 2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of medv') + theme(plot.title = element_text(hjust = 0.5)) 

qq_plot <- ggplot(crime_train, aes(sample=medv)) + stat_qq_point(color='dodgerblue') + stat_qq_line(color='darkgray') +
  labs(x="Thoretical Quantiles", y="Sample Quantiles", title = "QQ Plot of medv") + theme_classic() + 
  theme(plot.title = element_text(hjust = 0.5)) 

box_plot <- ggplot(crime_train, aes(x="", medv)) + geom_boxplot(fill='dodgerblue', color='darkgray')+ theme_classic() +
  labs(title = 'Boxplot of medv', x="") + theme(plot.title = element_text(hjust = 0.5)) + coord_flip()

box_target <- ggplot(crime_train, aes(x=factor(target), medv)) + geom_boxplot(fill='dodgerblue', color='darkgrey') +
  labs(x='target', title = 'Boxplot of medv by target') + theme_classic() +
  theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist, qq_plot, box_plot, box_target, ncol=2)
```

Finding correlations: The correlation plot below shows how variables in the dataset are related to each other. Looking at the plot, we can see that certain variables are more related than others.

```{r}
names(crime_train)
cor(drop_na(crime_train))
```

```{r}
pairs.panels(crime_train[1:14]) 
```

#DATA PREPARATION:

a. Missing Values - there are no missing values, so we will not do any missing value treatment.

b. outliers: I think we dont have any outliers that we should be removing at this stage.

c. Transformation -

age and lstat are both skewed, so lets see boxcox transformation suggestions.

```{r}
boxcoxfit(crime_train$age)
boxcoxfit(crime_train$lstat)
```

so for age the boxcox fit suggested power transformation of 1.3 and for lstat boxcox fit suggested power transformation of 0.23. Lets apply the same.

```{r}
crime_train$age_mod <- crime_train$age^1.3
crime_train$lstat_mod <- crime_train$lstat^0.23
```

The predictor dis, rm and medv has a moderate positive skew. Let's transform using the box-cox transformation

```{r}
boxcoxfit(crime_train$dis)
boxcoxfit(crime_train$rm)
boxcoxfit(crime_train$medv)
```

so for medv and rm the boxcox fit suggested power transformation of .23. Lets apply the same

```{r}
crime_train$medv_mod <- crime_train$medv^0.23
crime_train$rm_mod <- crime_train$rm^0.23
```

The lamda for the boxcoxfit for is dis is alose to 0, so we can apply log transformation.

```{r}
crime_train$dis_mod <- log(crime_train$dis)
```

Lets plot to see the status of the variables after transformation:

```{r}
hist_medv <- ggplot(crime_train, aes(medv_mod)) + geom_histogram(fill = 'dodgerblue', binwidth = 0.2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of medv_mod') + theme(plot.title = element_text(hjust = 0.5)) 

hist_lstat <- ggplot(crime_train, aes(lstat_mod)) + geom_histogram(fill = 'dodgerblue', binwidth = 0.2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of lstat_mod') + theme(plot.title = element_text(hjust = 0.5)) 

hist_dis <- ggplot(crime_train, aes(dis_mod)) + geom_histogram(fill = 'dodgerblue', binwidth = 0.2, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of dis_mod') + theme(plot.title = element_text(hjust = 0.5))

hist_rm <- ggplot(crime_train, aes(rm_mod)) + geom_histogram(fill = 'dodgerblue', binwidth = 0.025, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of rm_mod') + theme(plot.title = element_text(hjust = 0.5)) 

hist_age <- ggplot(crime_train, aes(age_mod)) + geom_histogram(fill = 'dodgerblue', binwidth = 50, color = 'darkgray' ) + 
 theme_classic() + labs(title = 'Histogram of age_mod') + theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(hist_medv, hist_lstat, hist_dis, hist_rm , hist_age, ncol=2)
```

We can see that the skewness of the transformed variables improved.

#BUILD MODELS:

Model 1 - : All original variables model . In this model we will use all the variables. This can be our base model and this model will not include any transformations. We can see which variables are significant. This will help us in looking at the P-Values and removing the non significant variables.

```{r}
model1 <- glm(target ~ zn + indus + chas + nox + rm + age + dis + rad + tax + ptratio + black + lstat + medv , family="binomial", data=crime_train) 
summary(model1)
```

Model 2: - All significant original variables model. I came up with this models after analyzing the output of model1. I removed all the variables that are not significant after seeing their P-Value.

```{r}
model2 <- glm(target ~ nox + age + dis + rad + tax + ptratio + black +  medv , family="binomial", data=crime_train) 
summary(model2)
```

Model 3: - All variables with transformations(will keep variables that were not transformed)

Model 3 includes original variables, plus the transformed variables from the transformations like power transformation and log transformations. This transfornation should help in reducing the skewness in the data or help them to become more normalized. This will help us in looking at the P-Values and removing the non significant variables.

```{r}
model3 <- glm(target ~ zn + indus + chas + nox + rm_mod + age_mod + dis_mod + rad + tax + ptratio + black + lstat_mod + medv_mod , family="binomial", data=crime_train) 
summary(model3)
```

Model 4: - Only the significant variables from model3 are used in this model. I removed all the variables that are not significant after seeing their P-Value.

```{r}
model4 <- glm(target ~ nox + age_mod + dis_mod + rad + ptratio + medv_mod , family="binomial", data=crime_train) 
summary(model4)
```

#MODEL SELECTION:

I would like to select Model3. The AIC and residual deviance for this model seemed to give the best values that would be suited for the prediction. Below is the ROC curve for model3 and to me it looks good. So i would like to proceed with model3

```{r}
crime_train$predict <- predict(model3, crime_train, type='response')

roc_model3 <- roc(crime_train$target, crime_train$predict, plot=T, asp=NA,
                legacy.axes=T, main = "ROC Curve", col="blue")

roc_model3["auc"]
```

Now lets do the confusion matrix:

```{r}
crime_train$predict_target <- ifelse(crime_train$predict >=0.5, 1, 0)
crime_train$predict_target <- as.integer(crime_train$predict_target)
myvars <- c("target", "predict_target")
crime_train_cm <- crime_train[myvars]
cm <- table(crime_train_cm$predict_target,crime_train_cm$target)
knitr:: kable(cm)
```

```{r}
Accuracy <- function(data) {
tb <- table(crime_train_cm$predict_target,crime_train_cm$target)
TN=tb[1,1]
TP=tb[2,2]
FN=tb[2,1]
FP=tb[1,2]
return((TP+TN)/(TP+FP+TN+FN))
}
Accuracy(data)
``` 

```{r}
CER <- function(data) {
tb <- table(crime_train_cm$predict_target,crime_train_cm$target)
TN=tb[1,1]
TP=tb[2,2]
FN=tb[2,1]
FP=tb[1,2]
return((FP+FN)/(TP+FP+TN+FN))
}
CER(data)
```

```{r}
Precision <- function(data) {
tb <- table(crime_train_cm$predict_target,crime_train_cm$target)
TP=tb[2,2]
FP=tb[1,2]
return((TP)/(TP+FP))
}
Precision(data)
```

```{r}
Sensitivity <- function(data) {
tb <- table(crime_train_cm$predict_target,crime_train_cm$target)
TP=tb[2,2]
FN=tb[2,1]
return((TP)/(TP+FN))
}
Sensitivity(data)
```

```{r}
Specificity <- function(data) {
tb <- table(crime_train_cm$predict_target,crime_train_cm$target)
TN=tb[1,1]
TP=tb[2,2]
FN=tb[2,1]
FP=tb[1,2]
return((TN)/(TN+FP))
}
Specificity(data)
```

```{r}
F1_score <- function(data) {
tb <- table(crime_train_cm$predict_target,crime_train_cm$target)
TN=tb[1,1]
TP=tb[2,2]
FN=tb[2,1]
FP=tb[1,2]
Precision = (TP)/(TP+FP)
Sensitivity = (TP)/(TP+FN)
Precision =(TP)/(TP+FP)
return((2*Precision*Sensitivity)/(Precision+Sensitivity))
}
F1_score(data)
```

#TEST DATA PREPARATION AND TESTING THE MODEL ON EVALUATION DATA:

In the final step we will test our model by using the test data.

```{r}
crime_eval <- read.csv("https://raw.githubusercontent.com/Riteshlohiya/Data621-Week3-Assignment3/master/crime-evaluation-data.csv")
crime_eval$age_mod <- crime_eval$age^1.3
crime_eval$lstat_mod <- crime_eval$lstat^0.23
crime_eval$dis_mod <- log(crime_eval$dis)
crime_eval$medv_mod <- crime_eval$medv^0.23
crime_eval$rm_mod <- crime_eval$rm^0.23
crime_eval$predict_prob <- predict(model3, crime_eval, type='response')
crime_eval$predict_target <- ifelse(crime_eval$predict_prob >= 0.50, 1,0)
write.csv(crime_eval,"Evaluation_Data.csv", row.names=FALSE)
```



























